Replacing handwritten signatures with open electronic signature software

### 2020-05-07 paper/main.tex
Fix last issues before submission to SVK

@@

\abstrakt{
So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako plnohodnotnú náhradu pre vlastnoručné podpisy.
Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.
V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
Výsledkom je otvorený, používateľsky prívetivý softvér pre platformy Windows, Mac OS X a Linux.
Pomocou aplikácie je možné, v súlade s nariadením eIDAS, podpísať a overiť podpis ľubovoľného dokumentu, ktorý má rovnaký právny účinok ako osvedčený podpis.
Naše riešenie podporuje občianske preukazy Slovenskej Republiky s čipom eID bez akýchkoľvek úprav alebo nastavení, čím sa umožní ich jednoduché využitie mimo obvyklého použitia vo webových aplikáciách štátu.
Aplikácia nie je viazaná iba na slovenské eID, ale je pripravená na ďalšie použitie s rôznymi zariadeniami v rôznych štátoch a ako platforma pre iné druhy podpisov.
Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch, aby bol projekt čo najviac prístupný verejnosti.
Dôraz kladieme na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
}

@@

There are several preconditions that allow us to consider electronic signatures as an alternative.
Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like \pojem{asymmetric cryptography} or \pojem{hashing}, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
We will concentrate on the first two in relation to the electronic signatures and where applicable open-source desktop computer software.

@@

With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \pojem{Act No. 215/2002 Coll.}. This situation changed with the EU Regulation eIDAS.

@@

\pojem{Qualified electronic signatures} (QES) are a subset of \pojem{advanced electronic signatures} with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
Qualified, in this case, means authorized for such use by the legal authorities.
These devices, certificates, and trust services are explicitly listed by the EU. In many cases indirectly by listing the organization (and its items) from a specific country.

@@

Such right can be exercised to license the software under either \pojem{proprietary} or \pojem{open-source} license.
\iffalse % Not part of the version for ŠVK
\pojem{Proprietary} meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
\pojem{Open-source} meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").

Motivation to license the software under an \pojem{open-source} license can differ, and so do such licenses.
While a \pojem{proprietary} license is usually made specifically for that entity and its interests, \pojem{open-source} licenses tend to be reused between different authors.
This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
\fi
In general, we can classify the \pojem{open-source} licenses into two categories: \pojem{permissive} and \pojem{copyleft}.

\pojem{Copyleft} license, in general, requires the user to publish the modified work under the compatible (free) license.
As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
\iffalse % Not part of the version for ŠVK
A popular example of such license is the GNU General Public License(GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
\fi

\pojem{Permissive} licenses, on the other hand, do not have such a requirement in place and are more suitable for potential commercialization.
They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
\iffalse % Not part of the version for ŠVK
Often used permissive licenses are MIT License, Apache License, or BSD License.
\fi

It is not uncommon to see \pojem{hybrid licensing} - licensing under more than one license, with one being typically \pojem{open-source} and one \pojem{proprietary}.
In this scenario, users can choose which license they want to use based on their needs.

@@

Previously mentioned standards \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES} define how \pojem{asymmetric cryptography}, \pojem{hashing}, or \pojem{timestamping} should be used to comply with the standards.
We go over a these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.

@@

These issues can be addressed with \pojem{asymmetric cryptography}.
In general, we assign a pair of keys to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
One party can encrypt, and the other is the only one that can decrypt.

\pojem{Digital signatures} are specific \pojem{asymmetric cryptography} algorithms.
The signer is the only one who possesses the \pojem{private key} used for signing, and others can use the \pojem{public key} for signature verification.
We sign the document, including the other parties' signature.

@@

We are left with the problem of public key distribution - even though \pojem{public-key schemes} do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.

A common solution is the use of \pojem{certificates}. \pojem{Certificates} are essentially \pojem{digital signatures} with metadata that can be used to establish the validity of the received \pojem{public key} before it is used.
Distribution of such \pojem{certificates} is the responsibility of a \pojem{Certification Authority} (CA) - a third party that all users in the network trust.
In the case of the Slovak ID cards, it is Disig - SVK eID Accredited CA, issued by the National Security Authority\footnote{Translation from "Národný bezpečnostný úrad".}.
Such \pojem{certificates}, therefore, reliably uniquely link and identify the \pojem{public key} owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
Ability to do that is the first requirement for the advanced electronic signatures mentioned in Section \ref{eidas}.
Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.

@@

\subsection{Hashing}

@@

In cryptography, we are considering only \pojem{hash functions} that can also satisfy these additional requirements:

\begin{enumerate}
  \item \textbf{Deterministic} - same input data are always mapped to the same \pojem{hash}.
  \item \textbf{One-way} - it should be infeasible to retrieve input data for a given \pojem{hash}.
  \item \textbf{Resistant to collisions} - it should be infeasible to find two inputs with the same \pojem{hash}.
  \item \textbf{Avalanche effect} - even small change to the input should lead to significantly different \pojem{hash}.
\end{enumerate}

In the context of \pojem{digital documents}, we can use such hash function to generate a hash for the document.
This hash can represent a document of any size and is used during the creation of a digital signature instead of the document itself.
Any further changes to the document will, therefore, lead to a significantly different hash, which means the digital signature will no longer be valid, and these changes are detected.
This ability is one of the requirements for the advanced electronic signatures from Section \ref{eidas}.
For this purpose, only hash functions that are reasonably fast can be used, so that we can utilize them on documents of variable size.

@@

\subsection{Timestamping}

@@

To achieve this, a trusted third party - \pojem{Timestamping Authority} (TSA) - is the one that will sign the \pojem{hash} of the document together with the current date and time.
This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
Verifying can be done by decrypting this information using the public key and comparing the hashes.
Any attempt at tampering is therefore detected, which helps to cover requirements for the advanced electronic signatures from Section \ref{eidas}.

@@

Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES}, respectively.
They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.

@@

  \caption{Overview of standards PAdES, XAdES, and CAdES}
  \label{table:1}
\end{table}

@@

@@

\obrazok{./figures/screenshot-jsignpdf}{Screenshot of the JSignPdf version 1.6.4}{screenshot-jsignpdf}

@@

\obrazok{./figures/screenshot-zepdisigsk}{Screenshot of the zep.disig.sk}{screenshot-zepdisigsk}

@@

\obrazok{./figures/screenshot-dpdfsigner}{Screenshot of the D.PDF Signer}{screenshot-dpdfsigner}

@@

\section{Overview}

  \caption{Comparison of existing software}
  \label{table:software}
\end{table}

@@

\obrazok{./figures/modularity}{Modules of our application and their communication}{modularity}

@@

\subsection{Downloads}

@@

\subsection{Help and resources}

@@

We proposed a simple software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
Our application was translated into three languages: English, Slovak, and Czech.
Our solution is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature where we take into consideration the velocity.
Our software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
From the software engineering perspective, we also covered testing and automation and how it can be used to correctly test and maintain Electron applications like ours in practice.
We compared our application with the existing software and found that there is room for improvement for our application in flexibility and ease of setup.
We also created a supporting website that provides users with downloads for all major desktop platforms, help, and general information on electronic signatures.

>>>>

\hyphenation{elektro-nic-kých}
\hyphenation{na-ria-dením}

@@

\abstrakt{
So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako plnohodnotnú náhradu pre vlastnoručné podpisy.
Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.\\
V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
Výsledkom je otvorený, používateľsky prívetivý softvér pre platformy Windows, Mac OS X a Linux.
Pomocou aplikácie je možné, v súlade s nariadením eIDAS, podpísať a overiť podpis ľubovoľného dokumentu, ktorý má rovnaký právny účinok ako osvedčený podpis.
Naše riešenie podporuje občianske preukazy Slovenskej Republiky s čipom eID bez akýchkoľvek úprav alebo nastavení, čím sa umožní ich jednoduché využitie mimo obvyklého použitia vo webových aplikáciách štátu.
Aplikácia nie je viazaná iba na slovenské eID, ale je pripravená na ďalšie použitie s rôznymi zariadeniami v rôznych štátoch a ako platforma pre iné druhy podpisov.
Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch, aby bol projekt čo najviac prístupný verejnosti.
Dôraz kladieme na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
}

@@

There are several preconditions that allow us to consider electronic signatures as an alternative.
Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like asymmetric cryptography or hashing, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
We will concentrate on the first two in relation to the electronic signatures and where applicable open-source desktop computer software.

@@

With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \textit{Act No. 215/2002 Coll.}. This situation changed with the EU Regulation eIDAS.

@@

\pojem{Qualified electronic signatures} (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
Qualified, in this case, means authorized for such use by the legal authorities.
These devices, certificates, and trust services are explicitly listed by the EU. In many cases indirectly by listing the organization (and its items) from a specific country.

@@

Such right can be exercised to license the software under either proprietary or open-source license.
\iffalse % Not part of the version for ŠVK
Proprietary meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
Open-source meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").

Motivation to license the software under an open-source license can differ, and so do such licenses.
While a proprietary license is usually made specifically for that entity and its interests, open-source licenses tend to be reused between different authors.
This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
\fi
In general, we can classify the open-source licenses into two categories: \pojem{permissive} and \pojem{copyleft}.

Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
\iffalse % Not part of the version for ŠVK
A popular example of such license is the GNU General Public License(GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
\fi

Permissive licenses, on the other hand, do not have such a requirement in place and are more suitable for potential commercialization.
They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
\iffalse % Not part of the version for ŠVK
Often used permissive licenses are MIT License, Apache License, or BSD License.
\fi

It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
In this scenario, users can choose which license they want to use based on their needs.

@@

Previously mentioned standards PAdES, XAdES, and CAdES define how asymmetric cryptography, hashing, or timestamping should be used to comply with the standards.
We go over a these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.

@@

These issues can be addressed with asymmetric cryptography.
In general, we assign a pair of keys to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
One party can encrypt, and the other is the only one that can decrypt.

\pojem{Digital signatures} are specific asymmetric cryptography algorithms.
The signer is the only one who possesses the private key used for signing, and others can use the public key for signature verification.
We sign the document, including the other parties' signature.

@@

We are left with the problem of public key distribution - even though public-key schemes do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.

A common solution is the use of \pojem{certificates}.
Certificates are essentially digital signatures with metadata that can be used to establish the validity of the received public key before it is used.
Distribution of such certificates is the responsibility of a \pojem{Certification Authority} (CA) - a third party that all users in the network trust.
In the case of the Slovak ID cards, it is Disig - SVK eID Accredited CA, issued by the National Security Authority\footnote{Translation from "Národný bezpečnostný úrad".}.
Such certificates, therefore, reliably uniquely link and identify the public key owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
Ability to do that is the first requirement for the advanced electronic signatures mentioned in Section \ref{eidas}.
Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.

@@

\subsection{Hashing and timestamping}

@@

In cryptography, we are considering only hash functions that can also satisfy these additional requirements:

\begin{enumerate}
  \item \textbf{Deterministic} - same input data are always mapped to the same hash.
  \item \textbf{One-way} - it should be infeasible to retrieve input data for a given hash.
  \item \textbf{Resistant to collisions} - it should be infeasible to find two inputs with the same hash.
  \item \textbf{Avalanche effect} - even small change to the input should lead to significantly different hash.
\end{enumerate}

In the context of digital documents, we can use such hash function to generate a hash for the document.
This hash can represent a document of any size and is used during the creation of a digital signature instead of the document itself.
Any further changes to the document will, therefore, lead to a significantly different hash, which means the digital signature will no longer be valid, and these changes are detected.
This ability is one of the requirements for the advanced electronic signatures from Section \ref{eidas}.
For this purpose, only hash functions that are reasonably fast can be used, so that we can utilize them on documents of variable size.

@@

@@

To achieve this, a trusted third party - \pojem{Timestamping Authority} (TSA) - is the one that will sign the hash of the document together with the current date and time.
This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
Verifying can be done by decrypting this information using the public key and comparing the hashes.
Any attempt at tampering is therefore detected, which helps to cover requirements for the advanced electronic signatures from Section \ref{eidas}.

@@

Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are PAdES, XAdES, and CAdES, respectively.
They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.

@@

  \caption{Overview of standards PAdES, XAdES, and CAdES.}
  \label{table:1}
\end{table}

@@

  \caption{Comparison of existing software.}
  \label{table:software}
\end{table}

@@

\obrazok{./figures/screenshot-jsignpdf}{Screenshot of the JSignPdf version 1.6.4.}{screenshot-jsignpdf}

@@

\obrazok{./figures/screenshot-zepdisigsk}{Screenshot of the zep.disig.sk.}{screenshot-zepdisigsk}

@@

\obrazok{./figures/screenshot-dpdfsigner}{Screenshot of the D.PDF Signer.}{screenshot-dpdfsigner}

@@

@@

\obrazok{./figures/modularity}{Modules of our application and their communication.}{modularity}

@@

@@

@@

Therefore, we proposed a simple open-source software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
Our application was translated into three languages: English, Slovak, and Czech.
It is also is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature.
The proposed software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
From the software engineering perspective, the implementation is robust, and we covered testing and automation in connection to Electron applications like ours.
On the other hand, there is some room for improvement in flexibility and ease of setup of our software.
We also maintain a supporting website that provides download links for all major desktop platforms, a help page, and general information on electronic signatures.
